Evaluating the Decency and Consistency of Data Validation Tests Generated by LLMs

### 2023-10-02 paper/paper.qmd
minor tweaks

---
title: "Evaluating the Decency and Consistency of Data Validation Tests Generated by LLMs"
author: 
    - name: Rohan Alexander^[Information and Statistical Sciences, University of Toronto. rohan.alexander@utoronto.ca.]
    - name: Lindsay Katz^[The Investigative Journalism Foundation]
    - name: Callandra Moore^[The Investigative Journalism Foundation]
    - name: Zane Schwartz^[The Investigative Journalism Foundation]
date: today
date-format: long
abstract: "We investigated the potential of large language models (LLMs) in developing dataset validation tests. We carried out 96 experiments each for both GPT-3.5 and GPT-4, examining different prompt scenarios, learning modes, temperature settings, and roles. The prompt scenarios were: 1) Asking for expectations, 2) Asking for expectations with a given context, 3) Asking for expectations after requesting a simulation, and 4) Asking for expectations with a provided data sample. For learning modes, we tested: 1) zero-shot, 2) one-shot, and 3) few-shot learning. We also tested four temperature settings: 0, 0.4, 0.6, and 1. Furthermore, two distinct roles were considered: 1) “helpful assistant”, 2) “expert data scientist”. To gauge consistency, every setup was tested five times. The LLM-generated responses were benchmarked against a gold standard suite, created by an experienced data scientist knowledgeable about the data in question. We find there are considerable returns to the use of few-shot learning, and that the more explicit the data setting can be the better. The results of the best combinations complement, but do not replace, those of the gold standard. The best LLM configurations complement, rather than substitute, the gold standard results. This study underscores the value LLMs can bring to the data cleaning and preparation stages of the data science workflow."
thanks: "We thank Monica Alexander. Code and data: https://github.com/RohanAlexander/evaluating_decency_and_consistency. Contributions: RA had the initial idea. RA and CM developed the experimental set-up. ZS established the political donations datasets. LK developed the initial suite of dataset validation tests that we compare the LLMs against. RA wrote the code to interact with the LLMs, evaluated the LLM outputs, and did the modelling. All authors contributed to writing the initial draft as well as improving and finalizing the paper."
bibliography: "bibliography.bib"
format: pdf
header-includes:
  - "\\usepackage{listings}"
  - "\\usepackage{fvextra}"
number-sections: true
---

@@

A human coder judges these responses produced by the LLMs, rating their decency (1-5, where 1 is the worst and 5 is the best) and their consistency (1-5, where 1 means they are very different and 5 means they are essentially identical). We then build an ordinal regression model with `rstanarm` to explore the relationships that decency and consistency have with prompts, roles, learning modes, and temperature settings.

We find there are considerable returns to one- and few-shot learning. We also find that including detailed descriptions of the expected dataset can help improve the quality of the tests that are produced, even more than including an example of the dataset. Surprisingly, there was not much of a difference found between GPT-3.5 and GPT-4.

Our results demonstrate one use for LLMs in a data analysis workflow, outside of just analysis. In particular, data scientists are often concerned that results may be an artifact of some errors in the data cleaning and preparation pipeline. Data validation tests help assuage these concerns, and using LLMs to establish an initial suite of tests can encourage their use.

@@

The Investigative Journalism Foundation (IJF) is a nonprofit news media outlet that is centered around public interest journalism. Their mandate is to help rebuild trust in Canadian democracy and hold the powerful accountable through data-driven investigative reporting. A core component of the IJF's work is building and actively maintaining eight publicly available databases with information on political donations, registered charities, and political lobbying in Canada. While the information that these databases are built upon are ostensibly public, they are not maintained or available in a way that is widely accessible or conducive to analysis. Further, the format and accessibility of these data vary greatly over time and across jurisdictions, making it incredibly difficult to look at temporal or regional differences in the data. In turn, the IJFs collation of these databases and high-quality journalism informed by these data serves as a crucial contribution to rebuilding trust and transparency in Canadian democracy.

One of the IJFs eight databases, and the focus of this work, is the political donations database. Canadian legislation requires political parties and candidates to disclose records of financial contributions they receive. These records are maintained by elections agencies across provinces and territories, and at the federal level. The frequency and scope of these disclosures vary across jurisdictions. For instance, in British Columbia, parties, candidates, constituency associations, and leadership and nomination contestants can receive political donations, while in Newfoundland and Labrador, donation recipients are limited to only parties and candidates [@ijfmethods]. The IJF's political donations database is a compilation of these political finance records across all Canadian jurisdictions, with data spanning from 1993 to the present day. The database contains 14 variables including the donor's name, the political party and entity to which the donation was made, the amount donated, as well as the region and year of the donation.

While the IJF's database is available in a clean, user-friendly format, the original records upon which it was created were not all accessible in this way. The format of donation records varies across jurisdiction and time. While some are available in readily downloadable spreadsheets, others are available as PDF or HTML files --- the former necessitating the use of optical character recognition (OCR) technology [@ijfmethods]. To prepare their database for publication, the IJF team performed significant manual cleaning. The majority of this work resulted from the conversion of PDF donation records to rectangular CSV format using OCR which is prone to scanning-related errors, such as the number 0 being scanned as the letter O. The IJF manually corrected these errors wherever they were identified by carefully comparing the machine-legible OCR output to the original static PDF donations record [@ijfmethods].

@@

Despite the thoughtful and thorough cleaning performed by the IJF team, the magnitude of these data coupled with their self-reported nature makes them prone to both human and computational error arising from the parsing process. With over 9.2 million rows in this database, it would be a massive undertaking for the IJF to manually identify all errors and inconsistencies present in their data --- whether that is as minor as an incorrectly formatted date, or as major as a reported donation amount thousands of dollars above the legal limit. To address this challenge, the IJF has been exploring the use of computational tools to assess data quality [@katzandmoore]. The team has developed a suite of comprehensive data tests for all eight databases using Python's `Great Expectations` library, which contains a number of pre-defined functions to test that particular characteristics expected of a dataset hold true in the data at large.

@@

Large language models (LLMs) are customized and refined through in-context learning using prompting [@ouyang2022]. A prompt is a set of natural language instructions which define the parameters and context for the desired output and may include one or more input-output examples. By using this approach, LLMs can apply their existing knowledge gained from training on various datasets to adapt to new contexts specified by the prompt. The outcomes generated by LLMs are highly sensitive to the specific phrasing and structure of these natural language instructions. Consequently, there is ongoing work to establish effective prompt patterns [@white2023].

@@

-   Four prompts:
```
        - The Investigative Journalism Foundation (IJF) created and maintains a CSV dataset related to political donations in Canada. Each observation in the dataset is a donation, and the dataset has the following variables: "index", "amount", "donor_location", "donation_date", "donor_full_name", "donor_type", "political_entity", "political_party", "recipient", "region", "donation_year", "amount_monetary", "amount_non_monetary", "electoral_event", "electoral_district", "added".\n\rPlease write a series of expectations using the Python package great_expectations for this dataset.
        - The Investigative Journalism Foundation (IJF) created and maintains a CSV dataset related to political donations in Canada. Each observation in the dataset is a donation, and the dataset has the following variables: "index", "amount", "donor_location", "donation_date", "donor_full_name", "donor_type", "political_entity", "political_party", "recipient", "region", "donation_year", "amount_monetary", "amount_non_monetary", "electoral_event", "electoral_district", "added". 
            - "amount" is a monetary value that cannot be less than $0. An example observation is "195.46". It is, possible, but unlikely to be more than $1,000.00. It cannot be NA. It should be a numeric. The maximum donation "amount" depends on the value of "region" and "year". For "Federal" is 1675, for "Quebec" is 100 since 2013 and 500 for earlier years, for "British Columbia" is 1309.09, for "Ontario" is 3325, and for "Alberta" is 4300. There is no limit for "Saskatchewan".
            - "amount" should be equal to the sum of "amount_monetary" and "amount_non_monetary".
            - "region" can be one of the following values: "Federal", "Quebec", "British Columbia", "Ontario", "Saskatchewan", "Alberta". It cannot be NA. It should be a factor variable.
            - "donor_full_name" is a string. It cannot be NA. It is usually a first and last name, but might also include a middle initial. It should be in title case.
            - "donation_date" should be a date in the following format: YYYY-MM-DD. It could be NA. The earliest donation is from 2010-01-01. The latest donation is from 2023-09-01.
            - "donation_year" should match the year of "donation_date" if "donation_date" is not NA, but it is possible that "donation_year" exists even if "donation_date" does not. The earliest year is 2010 and the latest year is 2023. This variable is an integer.
            - "political_party" cannot be NA. It should be a factor that is equal to one of: "New Democratic Party", "Liberal Party of Canada", "Conservative Party of Canada".

@@

            Based on this sample please write a series of expectations using the Python package great_expectations for this dataset.
```
-   Zero-, one-, and few-shot learning:
```
        - The following text in quotes is an example of an expectation for this dataset:
            """
            # Check that there is nothing null in any column of donations details
            donations_mv.expect_column_values_to_not_be_null(column='donor_full_name')
            """
        - The following text in quotes is an example of three expectations for this dataset:
            """
            # Check that there is nothing null in any column of donations details
            donations_mv.expect_column_values_to_not_be_null(column='donor_full_name')
            # Check that the federal donation does not exceed the maximum
            donations_mv.expect_column_values_to_be_between(
                column = 'amount',
                max_value = 1675,
                row_condition = 'region=="Federal" & donor_full_name.str.contains("Contributions Of")==False & donor_full_name.str.contains("Estate Of")==False & donor_full_name.str.contains("Total Anonymous Contributions")==False & donation_year == 2022 & political_entity.str.contains("Leadership")==False',
                condition_parser = 'pandas'
            )
            # Check that the date matches an appropriate regex format
            donations_mv.expect_column_values_to_match_regex(column = 'donation_date',
                                                        regex = '\\d{4}-\\d{2}-\\d{2}',
                                                        row_condition = "donation_date.isna()==False",
                                                        condition_parser = 'pandas')
            """
```
-   Four temperature values: 0, 0.4, 0.6, and 1.
-   Two roles:
```
        - You are a helpful assistant.
        - You are a highly-trained, experienced, data scientist who is an expert at writing readable, correct, Python code.
```

@@

There appears to be considerable difference in terms of the how the prompts are associated with decency. In particular, "Name" is never associated with a 5/5 rating (@fig-prompt-1). Surprisingly, however, the most informative prompt, "Example", is also never associated with a 5/5 rating. Instead it is "Describe" and "Simulate" that tend to be associated with better decency ratings. This is reflected in the averages, which are 2.65, 3.46, 3.44, and 2.94 for "Name", "Describe", "Simulate", and "Example", respectively.

@@

In contrast, as expected temperature has an effect on consistency. Temperature values of 0 are very clearly associated with ratings of high consistency, and higher temperatures are clearly associated with lower consistency (@fig-temperature-2). Their means differ considerably, with 4.77, 3.75, 3.31, and 2.69 for temperature values of 0, 0.4, 0.6, and 1, respectively.

@@

Role is an aspect of a prompt that is provided to the LLM before the main prompt content. We used two different roles, one that positioned the LLM as a helpful assistant, and the other that positioned the LLM as an experienced data scientist (@fig-role). We were expecting that the expert role would result in better code, but there was no obvious difference in terms of decency (@fig-role-1). There was also no obvious difference between the consistency (@fig-role-2). Their means did not differ by much in either case.

@@

Shots refers to the number of examples provided to the LLM as part of the prompt. Zero-shot means that no examples are provided, while one-shot and few-shot refer to one- and a few- examples being provided, respectively. Although the advantage of LLMs such as GPT-3.5 and GPT-4 is that they typically do well with zero-shot learning, we would expect that they will do better with one-shot and few-shot.

@@

Consistency and decency, our dependent variables, are ordered, categorical, outcomes. As such we model the rating $Y$, an ordinal outcome with $J = 5$ possible categories, using ordered logistic regression. Such a model works by estimating the probability that the outcome is less than or equal to some specific category, which in this case is the coded rank of that response (i.e. 1, 2, 3, 4, 5). Models for consistency and decency are estimated separately. In latent variable formulation, we assume that the observed category $y$ can be related to an underlying continuous latent variable, $y^*$, through a series of cutpoints:

@@

where $\zeta$ is a vector of cutpoints. The latent variable $y^*$ is then assumed to have a logistic distribution and is modeled as a linear function of covariates. In our case, we are interested in exploring the relationships that consistency and decency have with model, prompt, temperature, role, and number of shots:
$$
y^* = \beta_1 \cdot \mbox{version}_i + \beta_2 \cdot \mbox{prompt}_i + \beta_3 \cdot \mbox{temp}_i + \beta_4 \cdot \mbox{role}_i + \beta_5 \cdot \mbox{shot}_i
$$

In terms of our predictors, version is a binary variable as to whether we are using GPT-3.5 or GPT-4. We expect that GPT-4 will do better in terms of both decency and consistency than GPT-3.5. Prompt can be one of four values: "Name", "Describe", "Simulate", and "Example". We expect that "Simulate" and "Example" will be associated with higher decency than "Name" and "Describe". However, we expect the opposite relationship with consistency. This is because we expect that what will increase decency will be the more specific guidance provided by the "Simulate" and "Example" prompts, which should also increase the consistency. Temperature can take one of four values: 0, 0.4, 0.6 and 1. We expect that higher temperature values will be associated with less consistency. However, it is difficult to anticipate how temperature should be related to decency. Role can be one of two values: "Helpful", or "Expert", while shots can be one of three: "zero", "one", or "few". In the case of both role and shots, we expect that the "Expert" role, and one- and few-shot learning will be associated with higher decency, and consistency.

We fit this model, separately, for each of consistency and decency, in a Bayesian framework using the package `rstanarm` [@rstanarm] using the R statistical programming language [@citeR]. This computational process requires priors to be placed on $R^2$, the proportion of variance in the outcome that is attributable to the coefficients in a linear model, and the vector of cutpoints $\zeta$. For the prior on $R^2$, we follow @gelmanhillvehtari2020 [p. 276] and assume the mean is 0.3. For the vector cutpoints $\zeta$, we use an uninformative prior of $\text{Dirichlet}(1)$. @stanpolrvignette provides more information about fitting this type of model using `rstanarm`.

@@

```{r}
#| label: tbl-modelsummaryestimates
#| tbl-cap: "Exploring the relationships that consistency and decency have with model, prompt, temperature, role, and number of shots"
#| echo: false
#| message: false
#| warning: false

@@

```{r}
#| label: fig-modelsummaryestimates
#| fig-cap: "Exploring the relationships that consistency and decency have with model, prompt, temperature, role, and number of shots"
#| echo: false
#| message: false
#| warning: false

@@

@@

Our current experiments have been run in a very limited context --- namely one Canadian political donations dataset. Testing these prompts in a variety of contexts will provide evidence as to whether our prompting strategies are generalizable across datasets in different formats, contexts, and technical regimes. Prompts can be conceptualized as knowledge transfer methods analogous to software patterns which are intended to provide reusable solutions to common problems [@white2023]. This aligns with the goal of automated data validation to reduce the need for domain expertise and resources dedicated to data validation. Thus, in order to improve the explainability of our methods and to ensure our method is generalizable, we must validate that the prompts that are effective on this dataset show similar performance on other datasets from different contexts and in different formats.

>>>>

---
title: "Evaluating the Decency and Consistency of Data Validation Tests Generated by LLMs"
author: 
    - name: Rohan Alexander^[Information and Statistical Sciences, University of Toronto. rohan.alexander@utoronto.ca.]
    - name: Lindsay Katz^[The Investigative Journalism Foundation]
    - name: Callandra Moore^[The Investigative Journalism Foundation]
    - name: Zane Schwartz^[The Investigative Journalism Foundation]
date: today
date-format: long
abstract: "We investigated the potential of large language models (LLMs) in developing dataset validation tests. We carried out 96 experiments each for both GPT-3.5 and GPT-4, examining different prompt scenarios, learning modes, temperature settings, and roles. The prompt scenarios were: 1) Asking for expectations, 2) Asking for expectations with a given context, 3) Asking for expectations after requesting a simulation, and 4) Asking for expectations with a provided data sample. For learning modes, we tested: 1) zero-shot, 2) one-shot, and 3) few-shot learning. We also tested four temperature settings: 0, 0.4, 0.6, and 1. Furthermore, two distinct roles were considered: 1) "helpful assistant", 2) "expert data scientist". To gauge consistency, every setup was tested five times. The LLM-generated responses were benchmarked against a gold standard suite, created by an experienced data scientist knowledgeable about the data in question. We find there are considerable returns to the use of few-shot learning, and that the more explicit the data setting can be the better. The best LLM configurations complement, rather than substitute, the gold standard results. This study underscores the value LLMs can bring to the data cleaning and preparation stages of the data science workflow."
thanks: "We thank Monica Alexander and Chris Maddison for helpful suggestions. Code and data: https://github.com/RohanAlexander/evaluating_decency_and_consistency. Contributions: RA had the initial idea. RA and CM developed the experimental set-up. ZS established the political donations datasets. LK developed the initial suite of dataset validation tests that we compare the LLMs against. RA wrote the code to interact with the LLMs, evaluated the LLM outputs, and did the modelling. All authors contributed to writing the initial draft as well as improving and finalizing the paper."
bibliography: "bibliography.bib"
format:
  pdf:
    keep-tex: false
header-includes:
  - \usepackage{listings}
  - \lstset{basicstyle=\footnotesize\ttfamily,breaklines=true}
number-sections: true
---

@@

A human coder judges the responses produced by the LLMs, rating their decency (1-5, where 1 is the worst and 5 is the best) and their consistency (1-5, where 1 means they are very different and 5 means they are essentially identical). We then build an ordinal regression model with `rstanarm` to explore the relationships that decency and consistency have with prompts, roles, learning modes, and temperature settings.

We find there are considerable returns to one- and few-shot learning. We also find that including detailed descriptions of the expected dataset can help improve the quality of the tests that are produced, even more than including an example of the dataset. Temperature is associated with consistency with not with decency. Surprisingly, there was not much of a difference found between GPT-3.5 and GPT-4.

Our results demonstrate one use for LLMs in a data analysis workflow, outside of just analysis. In particular, data scientists are often concerned that results may be an artifact of some errors in the data cleaning and preparation pipeline. Data validation tests help assuage these concerns but they can be time consuming to produce. This paper shows it is possible to use LLMs to establish an initial suite of tests, which may encourage their use.

@@

The Investigative Journalism Foundation (IJF) is a nonprofit news media outlet that is centered around public interest journalism. Their mandate is to help rebuild trust in Canadian democracy and hold the powerful accountable through data-driven investigative reporting. A core component of the IJF's work is building and actively maintaining eight publicly available databases with information on political donations, registered charities, and political lobbying in Canada. While the information that these databases are built upon are ostensibly public, they are not maintained or available in a way that is widely accessible or conducive to analysis. Further, the format and accessibility of these data vary over time and across jurisdictions, making it difficult to look at temporal or regional differences in the data. In turn, the IJFs collation of these databases and high-quality journalism informed by these data serves as a crucial contribution to rebuilding trust and transparency in Canadian democracy.

One of the IJFs eight databases, and the focus of this work, is the political donations database. Canadian legislation requires political parties and candidates to disclose records of financial contributions they receive. These records are maintained by elections agencies across provinces and territories, and at the federal level. The frequency and scope of these disclosures vary across jurisdictions. For instance, in British Columbia, parties, candidates, constituency associations, and leadership and nomination contestants can receive political donations, while in Newfoundland and Labrador, donation recipients are limited to only parties and candidates [@ijfmethods]. The IJF's political donations database is a compilation of these political finance records across all Canadian jurisdictions, with data spanning from 1993 to the present day. The database contains 15 variables including the donor's name, the political party and entity to which the donation was made, the amount donated, as well as the region and year of the donation.

While the IJF's database is available in a clean, user-friendly format, the original records upon which it was created were not all accessible in this way. The format of donation records varies across jurisdiction and time. While some are available in readily downloadable spreadsheets, others are available as PDF or HTML files --- the former necessitating the use of optical character recognition (OCR) [@ijfmethods]. To prepare their database for publication, the IJF team performed significant manual cleaning. The majority of this work resulted from the conversion of PDF donation records to rectangular CSV format using OCR. This is prone to scanning-related errors, such as the number 0 being scanned as the letter O. The IJF manually corrected these errors wherever they were identified by comparing the machine-legible OCR output to the original PDF donations record [@ijfmethods].

@@

Despite the cleaning performed by the IJF team, the magnitude of these data coupled with their self-reported nature makes them prone to both human and computational error arising from the parsing process. With over 9.2 million rows in this database, it would be a massive undertaking for the IJF to manually identify all errors and inconsistencies present in their data --- whether that is as minor as an incorrectly formatted date, or as major as a reported donation amount thousands of dollars above the legal limit. To address this challenge, the IJF uses computational tools to assess data quality [@katzandmoore]. The team has developed a suite of comprehensive data tests for all eight databases using Python's `Great Expectations` library. This contains a number of pre-defined functions to test that particular characteristics expected of a dataset hold true in the data at large.

@@

Large language models (LLMs) are customized and refined through in-context learning using prompting [@ouyang2022]. A prompt is a set of natural language instructions which define the parameters and context for the desired output and may include one or more input-output examples. By using this approach, LLMs can apply their existing knowledge gained from training on various datasets to adapt to new contexts specified by the prompt. The outcomes generated by LLMs are sensitive to the specific phrasing and structure of these natural language instructions. Consequently, there is ongoing work to establish effective prompt patterns [@white2023].

@@

-   Four prompts:
\begin{lstlisting}
        - The Investigative Journalism Foundation (IJF) created and maintains a CSV dataset related to political donations in Canada. Each observation in the dataset is a donation, and the dataset has the following variables: "index", "amount", "donor_location", "donation_date", "donor_full_name", "donor_type", "political_entity", "political_party", "recipient", "region", "donation_year", "amount_monetary", "amount_non_monetary", "electoral_event", "electoral_district", "added".\n\rPlease write a series of expectations using the Python package great_expectations for this dataset.
        - The Investigative Journalism Foundation (IJF) created and maintains a CSV dataset related to political donations in Canada. Each observation in the dataset is a donation, and the dataset has the following variables: "index", "amount", "donor_location", "donation_date", "donor_full_name", "donor_type", "political_entity", "political_party", "recipient", "region", "donation_year", "amount_monetary", "amount_non_monetary", "electoral_event", "electoral_district", "added". 
            - "amount" is a monetary value that cannot be less than $0. An example observation is "195.46". It is, possible, but unlikely to be more than $1,000.00. It cannot be NA. It should be a numeric. The maximum donation "amount" depends on the value of "region" and "year". For "Federal" is 1675, for "Quebec" is 100 since 2013 and 500 for earlier years, for "British Columbia" is 1309.09, for "Ontario" is 3325, and for "Alberta" is 4300. There is no limit for "Saskatchewan".
            - "amount" should be equal to the sum of "amount_monetary" and "amount_non_monetary".
            - "region" can be one of the following values: "Federal", "Quebec", "British Columbia", "Ontario", "Saskatchewan", "Alberta". It cannot be NA. It should be a factor variable.
            - "donor_full_name" is a string. It cannot be NA. It is usually a first and last name, but might also include a middle initial. It should be in title case.
            - "donation_date" should be a date in the following format: YYYY-MM-DD. It could be NA. The earliest donation is from 2010-01-01. The latest donation is from 2023-09-01.
            - "donation_year" should match the year of "donation_date" if "donation_date" is not NA, but it is possible that "donation_year" exists even if "donation_date" does not. The earliest year is 2010 and the latest year is 2023. This variable is an integer.
            - "political_party" cannot be NA. It should be a factor that is equal to one of: "New Democratic Party", "Liberal Party of Canada", "Conservative Party of Canada".

@@

            Based on this sample please write a series of expectations using the Python package great_expectations for this dataset.
\end{lstlisting}
-   Three learning modes: zero-, one-, and few-shot learning:
\begin{lstlisting}
        - The following text in quotes is an example of an expectation for this dataset:
            """
            # Check that there is nothing null in any column of donations details
            donations_mv.expect_column_values_to_not_be_null(column='donor_full_name')
            """
        - The following text in quotes is an example of three expectations for this dataset:
            """
            # Check that there is nothing null in any column of donations details
            donations_mv.expect_column_values_to_not_be_null(column='donor_full_name')
            # Check that the federal donation does not exceed the maximum
            donations_mv.expect_column_values_to_be_between(
                column = 'amount',
                max_value = 1675,
                row_condition = 'region=="Federal" & donor_full_name.str.contains("Contributions Of")==False & donor_full_name.str.contains("Estate Of")==False & donor_full_name.str.contains("Total Anonymous Contributions")==False & donation_year == 2022 & political_entity.str.contains("Leadership")==False',
                condition_parser = 'pandas'
            )
            # Check that the date matches an appropriate regex format
            donations_mv.expect_column_values_to_match_regex(column = 'donation_date',
                                                        regex = '\\d{4}-\\d{2}-\\d{2}',
                                                        row_condition = "donation_date.isna()==False",
                                                        condition_parser = 'pandas')
            """
\end{lstlisting}
-   Four temperature values: 0, 0.4, 0.6, and 1.
-   Two roles:
\begin{lstlisting}
        - You are a helpful assistant.
        - You are a highly-trained, experienced, data scientist who is an expert at writing readable, correct, Python code.
\end{lstlisting}

@@

There appears to be considerable difference in terms of how the prompts are associated with decency. In particular, "Name" is never associated with a 5/5 rating (@fig-prompt-1). Surprisingly, however, the most informative prompt, "Example", is also never associated with a 5/5 rating. Instead it is "Describe" and "Simulate" that tend to be associated with better decency ratings. This is reflected in the averages, which are 2.65, 3.46, 3.44, and 2.94 for "Name", "Describe", "Simulate", and "Example", respectively.

@@

In contrast, as expected temperature has an effect on consistency. Temperature values of 0 are associated with ratings of high consistency, and higher temperatures are associated with lower consistency (@fig-temperature-2). Their means differ considerably, with 4.77, 3.75, 3.31, and 2.69 for temperature values of 0, 0.4, 0.6, and 1, respectively.

@@

Role is an aspect of a prompt that is provided to the LLM before the main prompt content. We considered two different roles, one that positioned the LLM as a helpful assistant, and the other that positioned the LLM as an experienced data scientist (@fig-role). We were expecting that the expert role would result in better code, but there was no obvious difference in terms of decency (@fig-role-1) or consistency (@fig-role-2). Their means did not differ by much in either case.

@@

Learning mode refers to the number of examples provided to the LLM as part of the prompt. Zero-shot learning means that no examples are provided, while one-shot and few-shot learning refer to one- and a few- examples being provided, respectively. Although the advantage of LLMs such as GPT-3.5 and GPT-4 is that they typically do well with zero-shot learning, we would expect that they will do better with one-shot and few-shot.

@@

Consistency and decency, our dependent variables, are ordered, categorical, outcomes. As such we model the rating $Y$, which is an ordinal outcome with $J = 5$ possible categories, using ordered logistic regression. Such a model works by estimating the probability that the outcome is less than or equal to some specific category, which in this case is the coded rank of that response (i.e. 1, 2, 3, 4, 5). Models for consistency and decency are estimated separately. In latent variable formulation, we assume that the observed category $y$ can be related to an underlying continuous latent variable, $y^*$, through a series of cutpoints:

@@

where $\zeta$ is a vector of cutpoints. The latent variable $y^*$ is then assumed to have a logistic distribution and is modeled as a linear function of covariates. In our case, we are interested in exploring the relationships that consistency and decency have with model, prompt, temperature, role, and learning mode:
$$
y^* = \beta_1 \cdot \mbox{version}_i + \beta_2 \cdot \mbox{prompt}_i + \beta_3 \cdot \mbox{temperature}_i + \beta_4 \cdot \mbox{role}_i + \beta_5 \cdot \mbox{shot}_i
$$

In terms of our predictors, version is a binary variable as to whether we are using GPT-3.5 or GPT-4. We expect that GPT-4 will do better in terms of both decency and consistency than GPT-3.5. Prompt can be one of four values: "Name", "Describe", "Simulate", and "Example". We expect that "Simulate" and "Example" will be associated with higher decency than "Name" and "Describe". However, we expect the opposite relationship with consistency. This is because we expect that what will increase decency will be the more specific guidance provided by the "Simulate" and "Example" prompts, which should also increase the consistency. Temperature can take one of four values: 0, 0.4, 0.6 and 1. We expect that higher temperature values will be associated with less consistency. However, it is difficult to anticipate how temperature should be related to decency. Role can be one of two values: "Helpful", or "Expert", while learning mode can be one of three: "zero", "one", or "few". In the case of both role and learning mode, we expect that the "Expert" role, and one- and few-shot learning will be associated with higher decency, and consistency.

We fit this model, separately, for each of consistency and decency, in a Bayesian framework using the package `rstanarm` [@rstanarm] and the R statistical programming language [@citeR]. This computational process requires priors to be placed on $R^2$, the proportion of variance in the outcome that is attributable to the coefficients in a linear model, and the vector of cutpoints $\zeta$. For the prior on $R^2$, we follow @gelmanhillvehtari2020 [p. 276] and assume the mean is 0.3. For the vector cutpoints $\zeta$, we use an uninformative prior of $\text{Dirichlet}(1)$. @stanpolrvignette provides more information about fitting this type of model using `rstanarm`.

@@

```{r}
#| label: tbl-modelsummaryestimates
#| tbl-cap: "Exploring the relationships that consistency and decency have with model, prompt, temperature, role, and learning mode"
#| echo: false
#| message: false
#| warning: false

@@

```{r}
#| label: fig-modelsummaryestimates
#| fig-cap: "Exploring the relationships that consistency and decency have with model, prompt, temperature, role, and learning mode"
#| echo: false
#| message: false
#| warning: false

@@

It is worth noting that the rating of the coherence and decency was a human endeavor, with all the bias that brings. In terms of decency, sometimes the LLMs simply did not generate code, instead generating natural language text about what tests could look like. Other times the LLMs generated a lot of very similar tests, for instance going through many possible options testing for class or non-null results, at the expense of developing a broadly useful suite. Finally, sometimes the LLMs would generate only one or two tests, or generate tests without any comments or other explanations. With regard to coherence, it was often the case that a low coherence rating was due to one of the five responses being especially difference. It was rare that the LLM just entirely misinterpreted the prompt, but compared with a human coder, who brings a broad context to writing tests, the LLMs rarely took into account the entire context that we would like.

@@

Our current experiments have been run in a very limited context --- namely one Canadian political donations dataset. We were also focused on writing tests within the `Great Expectations` framework.  It may be that testing these prompts in a variety of contexts, and also considering potential avenues of exploration beyond `Great Expectations`, will provide evidence as to whether our prompting strategies are generalizable across datasets in different formats, contexts, and technical regimes. Prompts can be conceptualized as knowledge transfer methods analogous to software patterns which are intended to provide reusable solutions to common problems [@white2023]. This aligns with the goal of automated data validation to reduce the need for domain expertise and resources dedicated to data validation. Thus, in order to improve the explainability of our methods and to ensure our method is generalizable, we must validate that the prompts that are effective on this dataset show similar performance on other datasets from different contexts and in different formats.
